\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-06-17 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

However, unlike the gradient drift case mentioned in Sec. \ref{subsec:GenGradDescent}, it is not possible to express the right hand side of \eqref{NewPDE} as the Wasserstein gradient of the functional \eqref{LyapButNotProx}.

@@

\appendices
\section{Showing $\Phi(\tilde{\rho})$ is Decreasing along the Flow Generated by \eqref{TransformedPDEivp}}\label{AppLyap}
Let $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ be the transient solution of \eqref{TransformedPDEivp}. Define
$$\bm{\ell}_{\bm{\xi}} := \frac{1}{2}\nabla_{\bm{\xi}}\log\tilde{\rho}, \quad \bm{\ell}_{\bm{\eta}} := \frac{1}{2}\nabla_{\bm{\eta}}\log\tilde{\rho},$$
and rewrite \eqref{NewPDE} as
$$\dfrac{\partial\tilde{\rho}}{\partial t} = - \nabla\cdot\left(\tilde{\rho}\bm{v}\right)$$
where $\nabla$ is the $2n\times 1$ Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and the vector field
$$\bm{v} := \begin{pmatrix}
 \bm{\eta}\\
 -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
 \end{pmatrix}
.$$

Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
\begin{align}
\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
\label{dPhidt}	
\end{align}
wherein the functional derivative
\begin{align}
\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
\label{FunctionalDerivative}
\end{align}
Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
\begin{strip}
\hrule
\begin{subequations}
\begin{align}
\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
\end{pmatrix}, \begin{pmatrix}
 \bm{\eta}\\
 -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
 \end{pmatrix}\bigg\rangle\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta}	\nonumber\\
&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
\end{align}
\label{ExplicitdPhidt}	
\end{subequations}
\hrule	
\end{strip}
where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.

>>>>

However, unlike the gradient drift case mentioned in Sec. \ref{subsec:GenGradDescent}, it is not possible to express the right hand side of \eqref{NewPDE} as the Wasserstein gradient of the functional \eqref{LyapButNotProx}. The latter is defined as \cite[Ch. 8]{ambrosio2008gradient} 
$$\nabla^{\text{Wasserstein}}\Phi := -\nabla\cdot\left(\rho\nabla\dfrac{\delta\Phi}{\delta\rho}\right),$$
where $\nabla$ denotes the standard Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and $\frac{\delta}{\delta\rho}$ denotes the functional derivative. From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$. Thus, even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, we cannot interpret it as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}.

@@

\appendices
\section{Showing $\Phi(\tilde{\rho})$ is Decreasing along the Flow Generated by \eqref{TransformedPDEivp}}\label{AppLyap}
Let $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ be the transient solution of \eqref{TransformedPDEivp}. Define
$$\bm{\ell}_{\bm{\xi}} := \frac{1}{2}\nabla_{\bm{\xi}}\log\tilde{\rho}, \quad \bm{\ell}_{\bm{\eta}} := \frac{1}{2}\nabla_{\bm{\eta}}\log\tilde{\rho},$$
and rewrite \eqref{NewPDE} as
$$\dfrac{\partial\tilde{\rho}}{\partial t} = - \nabla\cdot\left(\tilde{\rho}\bm{v}\right)$$
where $\nabla$ is the $2n\times 1$ Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and the vector field
\begin{align}
\bm{v} := \begin{pmatrix}
 \bm{\eta}\\
 -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
 \end{pmatrix}.
\label{defvectorfieldv}	
\end{align}

Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
\begin{align}
\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
\label{dPhidt}	
\end{align}
wherein the functional derivative
\begin{align}
\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
\label{FunctionalDerivative}
\end{align}
Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
\begin{strip}
\hrule
\begin{subequations}
\begin{align}
\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
\end{pmatrix}, \begin{pmatrix}
 \bm{\eta}\\
 -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
 \end{pmatrix}\bigg\rangle\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
\end{align}
\label{ExplicitdPhidt}	
\end{subequations}
\hrule	
\end{strip}
where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.
